Simulation Study

### 2018-03-23 datasets/RNA-Seq/yeast-simulation.Rmd

In this Rmd we will carry out the steps in `yeast48-benchmarkFDR.Rmd` for multiple
replicates, and plot results averaging over the replications.

@@

First, we download the processed count data from GitHub. There is one file for the 
Snf2 condition and one file for the wild type condition. The Snf2 condition is a yeast
strain that has the transcriptional regulator gene Snf2 knocked out. Each of these
files is a compressed `tar.gz` archive that contains a 
single bam file for each replicate in each condition, which we'll place in a subdirectory
called `data`.

@@

Here we'll carry out an analysis of the full dataset, comparing the controls to the
knowckout samples. The fold changes observed in this comparison will be used when 
generating the non-null simulated data in the following sections.

@@

There are many thousand genes detected using the full set. Next, we'll build input data frame for summarized benchmark.

>>>>

In this Rmd we will carry out simulations for multiple
replicates, and plot results averaging over the replications.

@@

First, we download the processed count data from GitHub. There is one file for the 
Snf2 condition and one file for the wild type condition. The Snf2 condition is a yeast
strain that has the transcriptional regulator gene Snf2 knocked out. Each of these
files is a compressed `tar.gz` archive that contains a 
single bam file for each replicate in each condition, which we'll place in a subdirectory
called `r datdir`.

@@

Here we'll carry out an analysis of the full dataset, comparing the controls to the
knockout samples. The fold changes observed in this comparison will be used when 
generating the non-null simulated data in the following sections.

@@

There are several thousand genes detected using the full set. Next, we'll build input data frame for summarized benchmark.

### 2018-03-23 datasets/simulations/parameter-simulation.Rmd

---
title: "Benchmarking FDR Methods - Simulation Summarized ROC / UpsetR Plots"
author: "Rafalab Journal Club Members"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    highlight: tango
    number_sections: true
    code_folding: hide
---

@@

<!-- ######################################################################## -->
## Workspace Setup
<!-- ######################################################################## -->

```{r, load-libraries, message=FALSE}
library("SummarizedBenchmark")
library("tidyverse")
```

```{r, helper-functions, message=FALSE}
source("../R/plotsim_aggupset.R")
```

In contrast to the t-test simulations, underlying data was not simulated in this series of simulations. Instead, for each simulation, a collection (20,000) of test statistics
were directly sampled from univariate distributions (Normal, non-central-t, or non-central-chi-squared). Test statistics corresponding to null tests are sampled with non-centrality
parameter zero, and non-null tests are sampled with some non-zero non-centrality parameter, determined based on the simulation setting. In these simulations, the magnitude of
the non-centrality parameter for each test can be understood as the effect size or signal in that test. The univariate distributions (Normal, t, chi-squared) are the variyng
sampling distributions that may be encountered under different testing scenarios (e.g. Student's t-test or Pearson's goodness-of-fit).

@@

<!-- ######################################################################## -->
## Local Helper Functions
<!-- ######################################################################## -->

```{r}
preclean <- function(res, merge_ihw = TRUE, supplementary = FALSE) {
    ## set of method names
    method_names <- unique(res$blabel)
    ## remove unadjusted p-values - too extreme
    if ("unadjusted" %in% method_names) {
        res <- res %>%
            filter(blabel != "unadjusted")
    }
    ## drop ihw not matching 'alpha'
    if (merge_ihw) {
        res <- res %>%
            mutate(ihw_alpha = gsub("ihw-a(.*)", "\\1", blabel),
                   ihw_alpha = ifelse(grepl("^ihw-a", blabel), ihw_alpha, alpha*100),
                   ihw_alpha = as.numeric(ihw_alpha) / 100) %>%
            filter(abs(alpha - ihw_alpha) < 1e-6) %>%
            mutate(blabel = ifelse(grepl("^ihw-a", blabel), "ihw", blabel))
    }
    ## drop excess Boca-Leek results
    if(!supplementary) {
        methods_bl <- grep("^bl-", method_names, value = TRUE)
        res <- res %>%
            filter(!(blabel %in% setdiff(methods_bl, "bl-df03")))
    }
    return(res)
}

myplot <- function(res, metrics = c("TPR", "FPR"), alpha_max = 1) {
    gp <- res %>%
        filter(performanceMetric %in% metrics) %>%
        filter (alpha <= alpha_max) %>%
        group_by(alpha, blabel, performanceMetric) %>%
        summarize(value50 = median(value, na.rm = TRUE),
                  value10 = quantile(value, 0.10, na.rm = TRUE),
                  value90 = quantile(value, 0.90, na.rm = TRUE)) %>%
        ungroup %>%
        ggplot(aes(x = alpha, y = value50, ymin = value10, ymax = value90,
                   group = blabel, color = blabel))
    if ("FPR" %in% metrics) {
        gp <- gp +
            geom_abline(aes(intercept = y, slope = s), alpha = 1/2,
                        data = data.frame(y = 0, s = 1, performanceMetric = "FPR"))
    }
    gp <- gp +
        geom_line(alpha = 1/4,
                  position = position_dodge(width = 0.003)) +
        geom_errorbar(width = .01,
                      position = position_dodge(width = 0.003)) + 
        geom_point(alpha = 1/2, position = position_dodge(width = 0.003)) +
        facet_wrap(~ performanceMetric) +
        scale_color_discrete("Method") + 
        ylab("performance across replicates") +
        xlab("alpha cutoff") +
        theme_bw()
    gp 
}
```

<!-- ######################################################################## -->
## Null Simulations
<!-- ######################################################################## -->

@@

### Normal

@@

### t (10 dof)

@@

### chi-sq (4 dof)

@@

<!-- ######################################################################## -->
## Informative Covariate Simulations
<!-- ######################################################################## -->

@@

<!-- ######################################################################## -->
## Uninformative Covariate Simulations
<!-- ######################################################################## -->

>>>>

---
title: "Simulation Study"
author: "Patrick Kimes"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
    html_document:
        toc: true
        toc_float: true
        highlight: tango
        number_sections: true
        code_folding: hide
---

@@

# Summary

For each simulation, a collection (20,000) of test statistics were directly
sampled from univariate distributions (Normal, non-central-t, or
non-central-chi-squared). Test statistics corresponding to null tests are
sampled with non-centrality parameter zero, and non-null tests are sampled
with some non-zero non-centrality parameter, determined based on the
simulation setting. In these simulations, the magnitude of the non-centrality
parameter for each test can be understood as the effect size or signal in
that test. The univariate distributions (Normal, t, chi-squared) are the variyng
sampling distributions that may be encountered under different testing scenarios
(e.g. Student's t-test or Pearson's goodness-of-fit).

@@

# Workspace Setup

```{r, wkspace-setup, results='hide', message=FALSE, warning=FALSE}
library("SummarizedBenchmark")
library("tidyverse")

source(file.path("..", "R", "plotsim_aggupset.R"))
```

# Helper Functions

```{r helper-preclean}
preclean <- function(res, merge_ihw = TRUE, supplementary = FALSE) {
    ## set of method names
    method_names <- unique(res$blabel)
    ## remove unadjusted p-values - too extreme
    if ("unadjusted" %in% method_names) {
        res <- res %>%
            filter(blabel != "unadjusted")
    }
    ## drop ihw not matching 'alpha'
    if (merge_ihw) {
        res <- res %>%
            mutate(ihw_alpha = gsub("ihw-a(.*)", "\\1", blabel),
                   ihw_alpha = ifelse(grepl("^ihw-a", blabel), ihw_alpha, alpha*100),
                   ihw_alpha = as.numeric(ihw_alpha) / 100) %>%
            filter(abs(alpha - ihw_alpha) < 1e-6) %>%
            mutate(blabel = ifelse(grepl("^ihw-a", blabel), "ihw", blabel))
    }
    ## drop excess Boca-Leek results
    if(!supplementary) {
        methods_bl <- grep("^bl-", method_names, value = TRUE)
        res <- res %>%
            filter(!(blabel %in% setdiff(methods_bl, "bl-df03")))
    }
    return(res)
}
```

```{r helper-myplot}
myplot <- function(res, metrics = c("TPR", "FPR"), alpha_max = 1) {
    gp <- res %>%
        filter(performanceMetric %in% metrics) %>%
        filter (alpha <= alpha_max) %>%
        group_by(alpha, blabel, performanceMetric) %>%
        summarize(value50 = median(value, na.rm = TRUE),
                  value10 = quantile(value, 0.10, na.rm = TRUE),
                  value90 = quantile(value, 0.90, na.rm = TRUE)) %>%
        ungroup %>%
        ggplot(aes(x = alpha, y = value50, ymin = value10, ymax = value90,
                   group = blabel, color = blabel))
    if ("FPR" %in% metrics) {
        gp <- gp +
            geom_abline(aes(intercept = y, slope = s), alpha = 1/2,
                        data = data.frame(y = 0, s = 1, performanceMetric = "FPR"))
    }
    gp <- gp +
        geom_line(alpha = 1/4,
                  position = position_dodge(width = 0.003)) +
        geom_errorbar(width = .01,
                      position = position_dodge(width = 0.003)) + 
        geom_point(alpha = 1/2, position = position_dodge(width = 0.003)) +
        facet_wrap(~ performanceMetric) +
        scale_color_discrete("Method") + 
        ylab("performance across replicates") +
        xlab("alpha cutoff") +
        theme_bw()
    gp 
}
```

# Null Simulations

@@

## Normal

@@

## t (10 dof)

@@

## chi-sq (4 dof)

@@

# Informative Covariate Simulations

@@

# Uninformative Covariate Simulations
